---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `wavfeatures`

<!-- badges: start -->
<!-- badges: end -->

The goal of `wavfeatures` is to extract features of a `.wav` file. For now, it is just the syllable nuclei.

## Installation

You can install the development version of wavfeatures like so:

``` r
remotes::install_github("tysonstanley/wavfeatures")
```

## Praat Script

The script the syllable nuclei is built on does the following:

Here's the step-by-step breakdown of what the script does:

### 1. Initial Setup:

Takes input parameters for:

* Ignorance Level/Intensity Median (0 or 2 dB)
* Minimum dip between peaks (2 or 4 dB)
* Directory path containing sound files

Processes all .wav files in the specified directory


### 2. For each sound file, it:

* Reads the sound file
* Removes DC offset (Subtract mean)
* Creates an intensity analysis (50ms window)

### 3. Peak Detection:

* Calculates noise floor (minimum intensity)
* Finds maximum intensity
* Calculates median intensity
* Sets a threshold = median intensity + ignorance level
* Identifies all intensity peaks above this threshold

### 4. Peak Validation:

* Checks each peak to ensure there's a sufficient dip in intensity before it
* Only keeps peaks where the difference between the peak and preceding dip is greater than the minimum dip parameter
* This helps distinguish between actual syllables and smaller intensity variations

### 5. Pitch Analysis:

* Performs pitch analysis to identify voiced segments
* Only keeps peaks that occur in voiced regions
* This helps eliminate false positives from noise or unvoiced consonants

### 6. TextGrid Creation:

* Creates a TextGrid file with two tiers: "syllables"
* Marks each validated, voiced peak as a point in the TextGrid
* Each point represents the center of a detected syllable
* Points are numbered sequentially (1, 2, 3, etc.)

### 7. The resulting TextGrid will contain:

* Point markers at each detected syllable nucleus
* Each point is labeled with a number indicating its sequence
* The points correspond to peaks in intensity that:
    a) Are above the threshold
    b) Have sufficient intensity dip before them
    c) Occur in voiced regions of speech

### Key considerations:

* The script may miss unstressed syllables (as noted in the comments)
* It works better on clean recordings (less background noise)
* Different settings for Ignorance Level and Minimum dip are recommended for filtered vs unfiltered sounds
* The accuracy depends on the quality of the recording and the speaker's articulation



## Example

This is a basic example:

```{r, eval=FALSE}
library(wavfeatures)

# Detect syllables in a WAV file
textgrid_path <- detect_syllables(
  "path/to/audio.wav",
  ignorance_level = 0,
  min_dip = 2
)

# Get the syllable count
syllable_count <- get_syllable_count(textgrid_path)
print(paste("Detected", syllable_count, "syllables"))
```
